---
title: "Eggcelent"
editor: visual
---

Let's make a simple predictor: 

- I've got some chickens at home, how many eggs will I get? 

```{r}
eggproduction  <- 
  readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2023/2023-04-11/egg-production.csv')
cagefreepercentages <-
  readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2023/2023-04-11/cage-free-percentages.csv')
```

## Setup Code

What is the average number of eggs per chicken? 

```{r}

```

If I have two chickens, how many eggs do I think I'll produce? 

```{r}

```

## Differences in Type 

What if I have some chickens that are free range, some that are organic *and* free range, and still others that are confinement raised? 

Now we need a model! 

```{r}

```














## Some Working Thoughts Below 






```{r}
eggproduction %>% 
  filter(prod_type == "hatching eggs") %>% 
  ggplot() + 
  aes(x=n_hens, y=n_eggs) + 
  geom_point()

eggproduction %>% 
  filter(prod_type == "table eggs") %>% 
  ggplot() + 
  aes(x=n_hens, y=n_eggs) + 
  geom_point()

eggproduction %>% 
  filter(prod_type == "table eggs") %>% 
  ggplot() + 
  aes(x=n_hens, y=n_eggs, color = prod_process) + 
  geom_point()

eggproduction %>% 
  filter(prod_type == "table eggs", prod_process == "all") %>% 
  ggplot() + 
  aes(x=n_hens, y=n_eggs) + 
  geom_point()

model <- 
  eggproduction %>% 
  filter(prod_type == "table eggs", prod_process == "all") %>% 
  lm(n_eggs ~ n_hens + I(n_hens^2), data = .)

df_augmented <- augment(model, eggproduction) %>% 
  filter(prod_type == "table eggs", prod_process == "all") %>% 
  ggplot() + 
  aes(x=n_hens, y=n_eggs) + 
  geom_point()

eggproduction %>% 
  ggplot() + 
  aes(
    x = n_hens, 
    y = n_eggs
  ) + 
  geom_point()

eggproduction %>% 
  ggplot() + 
  aes(x=n_hens, fill = paste0(prod_type, prod_process)) + 
  geom_histogram()

View(eggproduction)
```

